TGI-spinal-power

### 2022-05-18 analysis/tgispine_zoib_power-sim.Rmd
creating two power simulations - one for warm, one for cold

# Last edited: 17.05.2022

@@

# reshape data so that there is one rating column per participant, with a VAS column = type of rating
df_VAS <- melt(df, measure.vars= c("VAScold", "VASwarm","VASburn"), 
                        id.vars=c("ID","trial_n","manipulation","order",
                                  "trial_type","arm","condition","cold_location"),
             variable.name = 'VAS', value.name = 'rating')
# rename the VAS ratings
df_VAS <- df_VAS %>% 
  mutate(VAS = revalue(VAS, c("VAScold" = "cold",
                                    "VASwarm" = "warm",
                                    "VASburn" = "burn")))

@@

# Plotting data distributions to see what kind of model family is required
```{r}
# do this for the largest simulated dataset
df40$quality <- factor(df40$quality, 
                             levels = c("cold", "warm", "burn"))
ggplot2::ggplot(data = df40) +
  geom_density(aes(rating, fill = stimuli), alpha = .5) +
  facet_wrap(~quality) +
  theme_classic()
```
# Model function - Using zero inflated regression model
```{r}
# First model cold VAS
model <- function(x, vas = c('cold','warm','burn')) {
  if (vas[1] == 'cold'){
    x1 <- x %>%  filter(quality == 'cold')
  } else if (vas[1] == 'warm'){

@@

# Extract the power from each model
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  n = 40
  df40 <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df40 <- rbind(df40,df)}
  # then run the model based on simulated data
  m <- model(df40, 'warm')
  # extracting the p-value and effect size
  tmp <- c(m$coefficients$cond[4,4], m$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

@@

power <- sum(po$sig)/nSims
power
```

>>>>

# Last edited: 18.05.2022

@@

# reshape data so that there is one rating column per participant, with a VAS column = type of rating
df_VAS <- reshape2::melt(df, measure.vars= c("VAScold", "VASwarm","VASburn"), 
                        id.vars=c("ID","trial_n","manipulation","order",
                                  "trial_type","arm","condition","cold_location"),
             variable.name = 'VAS', value.name = 'rating')
# rename the VAS ratings
df_VAS <- df_VAS %>% 
  mutate(VAS = revalue(VAS, c("VAScold" = "cold",
                                    "VASwarm" = "warm",
                                    "VASburn" = "burn")))

@@

# Model function - Using zero inflated regression model
```{r}
# First model cold VAS
model <- function(x, vas = c('cold','warm','burn')) {
  if (vas[1] == 'cold'){
    x1 <- x %>%  filter(quality == 'cold')
  } else if (vas[1] == 'warm'){

@@

# Extract the power from each model - first cold
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
n <- 32 # number of participants to simulate
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  df_sim <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df_sim <- rbind(df_sim,df)}
  # then run the model based on simulated data
  m <- model(df_csim, 'cold')
  # extracting the p-value and effect size
  tmp <- c(m$coefficients$cond[4,4], m$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

@@

power <- sum(po$sig)/nSims # calculating power

# some renaming
po_c <- po
power_c <- power

power c
```

# Then warm
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
n <- 32 # number of participants to simulate
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  df_sim <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df_sim <- rbind(df_sim,df)}
  # then run the model based on simulated data
  m <- model(df_sim, 'warm')
  # extracting the p-value and effect size
  tmp <- c(m$coefficients$cond[4,4], m$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

power <- sum(po$sig)/nSims
# some renaming
po_w <- po
power_w <- power

power_w
```

# Plotting data distributions to see what kind of model family is required
```{r}
# do this for the largest simulated dataset
df40$quality <- factor(df40$quality, 
                             levels = c("cold", "warm", "burn"))
ggplot2::ggplot(data = df40) +
  geom_density(aes(rating, fill = stimuli), alpha = .5) +
  facet_wrap(~quality) +
  theme_classic()
```
